Silver Ion Chroma

### 2020-12-10 silver_ion_chroma.Rmd
added summary

---
title: "Silver Ion Chroma"
author: "Tristan Caro"
date: "12/10/2020"
output: html_document
---

A BAMES standard was run through two silver ion (Ag) columns to test the efficacy/cleanliness of the columns. A **Supelco Ag-Ion Discovery** column "Disc" and a homemade **Silver Nitrate** "Nitr" column were both tested (protocol courtesy of Nicolo Ardenghi).

@@

# Protocol
TC-BAME2 was taken from OGL Center Mini -20˚C. 245µg of total lipid BAME was suspended in 245µL of hex.
10µL (10µg) was removed and combined with 250µL hexane. The entirey of this stock (10µg) was applied to the Ag column.

0. Set up Ag-Ion columns over waste colleciton vials.
1. Condition SPE with 4mL acetone
2. Equilibrate cartridge with 4mL hexane. 
3. Load FAMEs in hexane derived from sample extraction at a flow rate of 2-3 drops/second. Remove waste collection vials and replace with fraction 1 (F1) collection vials (40mL glass vials).
4. Elute F1 with 6mL hexane:acetone 96:4
*Fraction 1 targets saturated FAs, trans monoenes, cis/cis and trans/trans conjugated linoleic acids*
5. Elute F2 with 4mL hexane:acetone 90:10
*Fraction 2 will target: cis monoenes, trans/trans dienes, CLAs*
6. Elute F3 with 4mL acetone
*Fraction 3 targets cis/cis dienes, other dienes, trienes.*
7. Evaporate all fractions at 40˚C under N2, sparge.
8. Reconstituted in three 100µL hexane washes, added to a GC vial with insert.
8. Evaporated at 40˚C under N2, reconstituted in 50µL cyclohexane for GC analysis.

@@

## Plot
```{r}
p_tc_bames2 <- results %>%
  filter(sample == c("tc_bames2")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  ggtitle("BAMES Standard 10µg/µL") +
  theme_bw()

p_tc_bames2 %>% ggplotly(dynamicTicks = TRUE)
```
```{r}
p_bames_disc_f1 <- results %>%
  filter(sample == c("bames_disc_f1")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Discovery Ag-Ion F1", subtitle = "saturated FAs, trans monoenes, cis/cis and trans/trans conjugated linoleic acids") +
  theme_bw()

p_bames_disc_f1 %>% ggplotly(dynamicTicks = TRUE)
```
```{r}
p_bames_disc_f2 <- results %>%
  filter(sample == c("bames_disc_f2")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Discovery Ag-Ion F2", subtitle = "cis monoenes, trans/trans dienes, CLAs") +
  theme_bw()

@@

```{r}
p_bames_disc_f3 <- results %>%
  filter(sample == c("bames_disc_f3")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Discovery Ag-Ion F3", subtitle = "cis/cis dienes, other dienes, trienes.") +
  theme_bw()

p_bames_disc_f3 %>% ggplotly(dynamicTicks = TRUE)
```
```{r}
p_bames_nitr_f1 <- results %>%
  filter(sample == c("bames_nitr_f1")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Ag - Nitrate F1") +
  theme_bw()

@@

```{r}
p_bames_nitr_f2 <- results %>%
  filter(sample == c("bames_nitr_f2")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Ag - Nitrate F2") +
  theme_bw()

p_bames_nitr_f2 %>% ggplotly(dynamicTicks = TRUE)
```
#### Issue with R plotting
Facet wrapping results in incomplete plotting - data points in tc_bames2 are not plotted.
```{r}
p_tc_bames2 <- results %>%
  filter(sample == c("tc_bames2", "bames_disc_f1", "bames_nitr_f1")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  facet_wrap(~sample, nrow = 1, scales = "free_y") +
  scale_fill_viridis_d() +
  ggtitle("BAMES Standard 10µg/µL") +
  theme_bw()

>>>>

---
title: "Silver Ion Chroma"
author: "Tristan Caro"
date: "12/10/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

# Summary
For those compounds observed in F1, the silver ion column resulted in an order-of-magnitude decrease in TIC. F1 was greatly depleted in hydroxy-bearing FAMES.

No compounds eluted in F2, which targets cis monoenes, trans/trans dienes, and CLAs. All monoenes eluted in F1.

None of the compounds expected in F3 eluted (dienes). Oddly, one compound eluted in F3 of the discovery column, FAME C16:0, which would be expected to elute in F1.

Silocone signal was observed across all samples. Silicone peaks coeluted with FAMEs 3-OH 14:0 (28 minutes) 17:0 (31 minutes), and 20:0 (36 minutes). My guess is that this is a product of the column production/cleaning.

```{r, echo = FALSE}
knitr::include_graphics('images/silocone_ms.png')
```

# Protocol
A BAMES standard was run through two silver ion (Ag) columns to test the efficacy/cleanliness of the columns. A **Supelco Ag-Ion Discovery** column "Disc" and a homemade **Silver Nitrate** "Nitr" column were both tested (protocol courtesy of Nicolo Ardenghi).

`TC-BAME2` was taken from OGL Center Mini -20˚C. 245µg of total lipid BAME was suspended in 245µL of hex.
10µL (10µg) was removed and combined with 250µL hexane. The entirey of this stock **(10µg)** was applied to the Ag column.

### Discovery Ag-Ion Column Chromatography
0. Set up Ag-Ion columns over waste colleciton vials.
1. Condition SPE with 4mL acetone
2. Equilibrate cartridge with 4mL hexane. 
3. Load FAMEs in hexane derived from sample extraction at a flow rate of 2-3 drops/second. Remove waste collection vials and replace with fraction 1 (F1) collection vials (40mL glass vials).
4. Elute F1 with 6mL hexane:acetone 96:4
*Fraction 1 targets saturated FAs, trans monoenes, cis/cis and trans/trans conjugated linoleic acids*
5. Elute F2 with 4mL hexane:acetone 90:10
*Fraction 2 will target: cis monoenes, trans/trans dienes, CLAs*
6. Elute F3 with 4mL acetone
*Fraction 3 targets cis/cis dienes, other dienes, trienes.*
7. Evaporate all fractions at 40˚C under N2, sparge.
8. Reconstituted in three 100µL hexane washes, added to a GC vial with insert.
8. Evaporated at 40˚C under N2, reconstituted in 50µL cyclohexane for GC analysis.

**GC-MS**

@@

```
Autosampler Position: Sample
37: Cyclohex Blank
38: TC-BAMES-2
39: BAMES F1 Disc
40: BAMES F2 Disc
41: BAMES F3 Disc
42: BAMES F1 Nitr
43: BAMES F2 Nitr
37: Cyclohex Blank
```

NOTE: F3 of Ag-Nitrate was disqualified due to the presence of Ag-nitrate powder in the sample collection vial.

@@

## Plot
```{r, echo = FALSE, warning=FALSE}
p_tc_bames2 <- results %>%
  filter(sample == c("tc_bames2")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  ggtitle("BAMES Standard 10µg/µL") +
  theme_bw()

p_tc_bames2 %>% ggplotly(dynamicTicks = TRUE)
```
```{r, echo = FALSE, warning=FALSE}
p_bames_disc_f1 <- results %>%
  filter(sample == c("bames_disc_f1")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Discovery Ag-Ion F1", subtitle = "saturated FAs, trans monoenes, cis/cis and trans/trans conjugated linoleic acids") +
  theme_bw()

p_bames_disc_f1 %>% ggplotly(dynamicTicks = TRUE)
```
```{r, echo = FALSE, warning=FALSE}
p_bames_disc_f2 <- results %>%
  filter(sample == c("bames_disc_f2")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Discovery Ag-Ion F2", subtitle = "cis monoenes, trans/trans dienes, CLAs") +
  theme_bw()

@@

```{r, echo = FALSE, warning=FALSE}
p_bames_disc_f3 <- results %>%
  filter(sample == c("bames_disc_f3")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Discovery Ag-Ion F3", subtitle = "cis/cis dienes, other dienes, trienes.") +
  theme_bw()

p_bames_disc_f3 %>% ggplotly(dynamicTicks = TRUE)
```
```{r, echo = FALSE, warning=FALSE}
p_bames_nitr_f1 <- results %>%
  filter(sample == c("bames_nitr_f1")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Ag - Nitrate F1") +
  theme_bw()

@@

```{r, echo = FALSE, warning=FALSE}
p_bames_nitr_f2 <- results %>%
  filter(sample == c("bames_nitr_f2")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  scale_fill_viridis_d() +
  labs(title = "BAMES Ag - Nitrate F2") +
  theme_bw()

p_bames_nitr_f2 %>% ggplotly(dynamicTicks = TRUE)
```

### Issue with R plotting
Facet wrapping results in incomplete plotting - data points in tc_bames2 are not plotted.
```{r}
p_tc_bames2 <- results %>%
  filter(sample == c("tc_bames2", "bames_disc_f1", "bames_nitr_f1")) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1)+
  facet_wrap(~sample, nrow = 1, scales = "free_y") +
  scale_fill_viridis_d() +
  ggtitle("BAMES Standard 10µg/µL") +
  theme_bw()
